Against the Planetary Interpretation of PTFO$\,$8-8695b

### 2020-04-03 paper/ms.tex
little bit of extra text

\section{Observations}
\label{sec:observations}

% \subsection{Observations}

PTFO 8-8695 was observed by TESS with Camera 1, CCD 1, from December
15, 2018 to January 6, 2019, during the sixth sector of science
operations.  The star is designated TIC 264461976 in the TESS Input
Catalog \citep{stassun_TIC_2018}.  The pixel data for an $11\times11$
array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
the onboard computer. 
Each 2048$\times$2048 image was also averaged into 30-minute stacks,
and saved as a ``full frame image'' (FFI).

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
The data were downlinked via the Deep Space
Network\footnote{\url{deepspace.jpl.nasa.gov}}, and the
spacecraft timestamps were calibrated against the ground-station
clocks.  The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve, which has had non-astrophysical variability removed
through the methods discussed by \citet{smith_kepler_apertures_2017}
and \citet{smith_kepler_PDC_2017}.  We then processed the lightcurve
as follows.  First, we removed all points with non-zero quality flags.
This removed data that might have been adversely affected by
``momentum dumps,'' the firing of thrusters and resetting of reaction
wheels\footnote{The spacecraft pointing and momentum dumps are
described in the data release notes:
\url{archive.stsci.edu/missions/tess/doc/tess_drn/tess_sector_07_drn09_v03.pdf}}
that took place every 2.5 days during sector 2.  The data during these
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude'' \citep[][Table
28]{tess_data_product_description_2018}.  For WASP-4, these flags were
simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Out of caution, we clipped out an additional 10 minutes before and
after every momentum dump.  We also removed the data within the first
and last hours of both orbits, because of correlated red noise that
appears during those time ranges.

%FIXME these paragraphs are directly copied from WASP-4 paperI
%FIXME these paragraphs are directly copied from WASP-4 paperI
All told, we removed 8\% of the original data points, and were left
with 18{,}165 measurements of the relative flux of WASP-4.  We
normalized the data by dividing out the median flux.  We converted the
timestamps from BTJD$_{\rm TDB}$ into BJD$_{\rm TDB}$ by adding the
appropriate 2{,}457{,}000 day offset
\citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
``flatten'' the lightcurves, as is often done with splines,
polynomials, or Gaussian processes.  Instead, we modeled the
out-of-transit flux variations simultaneously with the transit
parameters, as described below.

@@

\section{Analysis}
\label{sec:analysis}

@@

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}. 
The two largest peaks in the Lomb-Scargle periodogram of the
light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.
The $P_{\rm \ell}$ peak had the larger ampitude of the two.
Smaller harmonics from each of these dominant two signals were also
present.

Initial fitting experiments with splines showed that subtracting out
the long-period signal, the short-period signal dominated the
periodogram, and vice-versa.
However it quickly became clear that a mroe careful modelling approach
than ``flattening'' the light-curve with splines or Gaussian proceses
would be preferable, in order to preserve the power at each frequency.

@@

We opted to model the light-curve as a linear combination of Fourier
harmonics at the short and long periods, plus a transit at the
short period.
Symbolically,
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell} = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the relative flux at the short period, and
$f_{\rm \ell}$ is the flux at the long period.
Decomposing each term,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{n=N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{n=N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{m=M} B_m \cos([m\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
for $N$ and $M$ the total number of harmonics at the short and long period, respective,
$A_i$ and $B_i$ for each harmonic term, and $\omega_i = 2\pi / P_i$ the angular
frequency for $i$ the short or long period index. 
We fixed the ``phase-offset'' for the short period signal to be zero,
and let the reference time of the long period signal float by introducing
$\phi_{\rm \ell}$. 
Since we did not a priori know how many harmonics would be
appropriate, we let the number of harmonics vary, and
used the Bayesian information criterion to choose the
appropriate model.

For example, one possible model could be a transit, plus $N=2$ harmonics
of sines and cosines at the short period, plus $M=1$  harmonics at the long
period.
In this case, the free parameters would be as follows.
For the transit, we would fit for
the impact parameter, the planet-to-star radius ratio, two quadratic limb darkening parameters,
the planet orbital period (equal to the short period), the reference time
for the transit, and the mean flux.
There would be 4 additional Fourier amplitudes at the short period,
plus 2 Fourier amplitudes at the long period, and well as the long period itself
and its phase.
For this case, we therefore fitted 14 free parameters.

@@

\section{Discussion}
\label{sec:discussion}

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

>>>>

\section{The Data}
\label{sec:observations}

\subsection{Observations}

PTFO 8-8695 was observed by TESS with Camera 1, CCD 1, from December
15, 2018 to January 6, 2019, during the sixth sector of science
operations.  The star is designated TIC 264461976 in the TESS Input
Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.  The pixel data for an $11\times11$
array surrounding PTFO 8-8695 were averaged into 2-minute stacks by
the onboard computer. 
Each 2048$\times$2048 image from the CCD was also averaged into
30-minute stacks, and saved as a ``full frame image'' (FFI).

The 2-minute stacks for PTFO 8-8695 were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
Presearch Data Conditioning (PDC) lightcurve, which had
non-astrophysical variability removed through the methods discussed by
\citet{smith_kepler_apertures_2017} and \citet{smith_kepler_PDC_2017}.

We also processed the 30-minute image stacks separately, as part of
the Cluster Difference Imaging
Photometric Survey (CDIPS; \citep{bouma_cluster_2019}).
The raw CDIPS light-curve shows identical features as the 2-minute
light-curve, though at 30-minute cadence.

%FIXME: these numbers
%FIXME: gotta mask orbit edges too...
Both light-curves showed a clear discontinuous ``jump'' in the last
few days, which seemed likely to be an instrumental systematic.
We masked out HOW MUCH DATA(?).
As an added precaution, we masked out the first and last N hours
of each orbit, since there is often
systematic red noise during those times.
Finally, we removed all points with non-zero quality flags \citep[see][]{tess_data_product_description_2018}.

%FIXME: these numbers
All told, we removed N\% of the original data points, and were left
with NN{,}NNN measurements of the relative flux of PTFO 8-8695.  We
normalized the data by dividing out the median flux, and then
subtracted out the mean.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two largest peaks in the Lomb-Scargle periodogram of the
light-curve were clearly separated at $P_{\rm s} \approx 0.448\,{\rm
days}$ and $P_{\rm \ell} \approx 0.499\,{\rm days}$.  The $P_{\rm
\ell}$ peak had the greater power of the two.  Smaller harmonics from
each of these dominant two signals were also present.

% 0.14 = A1+A2
% 0.06 = A1-A2
% 2A1 = 0.20
% -> A1 = 0.1
% 2A2 = 0.08
% -> A2 = 0.04

The beat peak-to-peak amplitude at maximum, when the two signals
constructively interfere, is about 14\%.  At minimum, the peak-to-peak
amplitude is about 6\%.  Assuming the signals are just two sinusoids,
algebra tells us that the peak-to-peak amplitudes should therefore be
10\% for the long-period signal, and 4\% for the short-period signal.
These numbers will turn out to be over-estimates, but only slightly.

Initial signal-processing experiments fitting out splines or sinusoids
showed that after subtracting out the long-period signal, the short-period
signal dominated the periodogram, and vice-versa.
However it quickly became clear that it would be beneficial to simultaneously
model the signals separately, in order to
preserve the power at each frequency.

@@

\section{The Model}
\label{sec:analysis}

@@

@@

We opted to model the light-curve as a linear combination of Fourier
harmonics at the short and long periods, plus a transit at the short
period.  Symbolically,
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the relative flux at the short period, and
$f_{\rm \ell}$ is the flux at the long period.  Writing out the
Fourier terms,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{n=N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{n=N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{m=M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{m=M} B_m \cos([m\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
for $N$ and $M$ the total number of harmonics at the short and long
period, respective, $A_i$ and $B_i$ for each harmonic term, and
$\omega_i = 2\pi / P_i$ the angular frequency for $i$ the short or
long period index.  We fixed the ``phase-offset'' for the short period
signal to be zero, and let the reference time for the long period
signal float by introducing $\phi_{\rm \ell}$.  Since we did not a
priori know how many harmonics would be appropriate, we considered a
number of different choices for $N$ and $M$, and used the Bayesian
information criterion to choose the appropriate model.

As an example, one possible model could be a transit, plus $N=2$
harmonics of sines and cosines at the short period, plus $M=1$
harmonics at the long period.  In this case, the free parameters would
be as follows.  For the transit, we would fit for the impact
parameter, the planet-to-star radius ratio, two quadratic limb
darkening parameters, the planet orbital period (equal to the short
period), the reference time for the transit, and the mean flux.  There
would be 4 additional Fourier amplitudes at the short period, plus 2
Fourier amplitudes at the long period, and well as the long period
itself and its phase.  For this case, we therefore fitted 14 free
parameters.

We implemented the models using \texttt{PyMC3}, which is built on \texttt{theano} \citep{salvatier_2016_PyMC3,exoplanet:theano}.
For the Fourier terms, we used the default math operators.
For the exoplanet transit, we used the model and derivatives implemented in
\texttt{exoplanet} \citep{exoplanet:exoplanet}.

Our priors were as follows.
%FIXME

We initialized to the maximum a posteriori (MAP) solution.
We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turns Sampler, 
and used R-hat as our convergence diagnostic.

@@

\section{Interpretation}
\label{sec:discussion}

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}
